---
title: "Peer Assessment 1"
output: html_document
---

Part - 1: Loading and Pre-Processing Data
```{r}
library(ggplot2)
library(scales)
library(grid)
library(gridExtra)
```
```{r, echo=TRUE}
#loading of data from "activity.csv"
connection <- unz("activity.zip","activity.csv")
activity <- read.table(file = connection,header = TRUE,sep = ",")

#Preprocessing Data
activity$datetime <- as.POSIXct(with(activity,paste(date,paste(interval %/%100, interval %% 100, sep=":"))),format="%Y-%m-%d %H:%M",tz="")
```
End of Part - 1

Part - 2: Mean total number of steps taken each day.
```{r, echo=TRUE}
# Making the distribution.
stepsPerDay <- setNames(aggregate(steps~as.Date(date),activity,sum,na.rm =TRUE),c("date","steps"))
# plotting the histogram for the above distribution
hist1 <- ggplot(stepsPerDay,aes(x=date,y=steps))+geom_bar(stat="identity") + ggtitle("Total number of steps per day (source data)")
print(hist1)
```
```{r, echo=TRUE}
#Summary
summary(stepsPerDay$steps)
#Displaying Mean and Median
mm_1 <- c(mean = mean(stepsPerDay$steps),median = median(stepsPerDay$steps))
print(mm_1)
```

End of Part - 2.

Part - 3: Finding Average daily Activity Patteern.
Question 1:  
```{r, echo=TRUE}
avgPattern <- aggregate(steps~interval,activity,mean,na.rm = TRUE)
avgPattern$time <- as.POSIXct(with(avgPattern,paste(interval %/% 100, interval%% 100, sep=":")),format="%H:%M")
plotp3 <- ggplot(avgPattern,aes(x=time,y=steps)) + geom_line() + scale_x_datetime(breaks = date_breaks("2 hour"),labels = date_format("%H:%M"))
print(plotp3)
```
Question 2:
```{r, echo=TRUE}
with(avgPattern,avgPattern[steps == max(steps),])
```

End of Part - 3.

Part - 4: Imputing missing values.
```{r, echo=TRUE}
# Finding missing data..
mis <- aggregate(cnt~date,cbind(activity[is.na(activity$steps),],cnt=c(1)),sum,na.rm = FALSE)
mis$dow <- weekdays(as.Date(mis$date),abbreviate=TRUE)
print(mis[,c(1,3,2)])
unique(mis$dow)
```
Number of missing data values are 288*8 = 2,304.
```{r}
# Reference dataset for calculating the missing values.
ref_dataset <- aggregate(steps~interval+weekdays(datetime,abbreviate=TRUE),activity,FUN=mean,na.rm=TRUE)
colnames(ref_dataset) <- c("interval","dow","avg_steps")
ref_dataset$dow <- factor(ref_dataset$dow,levels = c("Mon","Tue","Wed","Thu","Fri","Sat","Sun"))
ggplot(ref_dataset,aes(x=interval,y=avg_steps)) + geom_line() + facet_grid("dow ~ .")
# Creating A fixed Dataset.
activity$dow <- weekdays(activity$datetime,abbreviate=TRUE)
actfin <- merge(activity,ref_dataset,by=c("dow","interval"),all.x = TRUE)
actfin <- actfin[with(actfin,order(date,interval)),]
actfin$fixed_steps <- ifelse(is.na(actfin$steps),actfin$avg_steps,actfin$steps)
```
Comparing Datasets
```{r, echo=TRUE}
#Making the dataset for Histogram
stepsPerDay2 <- setNames(aggregate(fixed_steps~as.Date(date),actfin,sum,na.rm = TRUE),c("date","steps"))
```
```{r, echo=TRUE}
# Plotting histogram
hist2 <- ggplot(stepsPerDay2,aes(x=date,y=steps)) + geom_bar(stat="identity")+ggtitle("Total number of steps per day (fixed data)")
```
**Combining view for comparison**
```{r, echo=TRUE}
grid.arrange(hist1, hist2, nrow=2)
```
```{r, echo=TRUE}
mm_2 <- c(mean = mean(stepsPerDay2$steps),median = median(stepsPerDay2$steps))
comparison <- rbind(source = mm_1, fixed = mm_2, delta = mm_2-mm_1)
print(comparison)
```

End of part - 4.

Part - 5: Checking for difference in activity of weekdays and weekends.
```{r, echo=TRUE}
week_diff <- aggregate(steps~dow+interval, with(activity,data.frame(dow = factor(ifelse(weekdays(as.Date(date)) %in% c("Sunday","Saturday"),"weekend","weekday" )),interval,steps)),FUN = mean,rm.na = TRUE) # group steps by weekend/weekday and interval to find average steps
ggplot(week_diff,aes(x=interval,y=steps)) + geom_line() + facet_grid("dow ~ .")
```
End of Part - 5.

End of Assignment.
=================